---
title: "Using flashr for mashr prior specification"
author: "Gao Wang and others"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{using flashr for mashr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,comment = "#",fig.width = 5,
                      fig.height = 4,fig.align = "center",
                      eval = TRUE)
```

# Introduction

This is continuation of the [eQTL analysis vignette][eqtl].
In that vignette we have used PCA to compute data driven covariances. Here we demonstrate
the use of additional data driven covariance, via [flash](https://github.com/stephenslab/flashr) decomposition.

# Dataset simulation

Same as the [eQTL analysis vignette][eqtl] we simulate a toy data-set,

```{r}
library(ashr)
library(mashr)
set.seed(1)
simdata = simple_sims(10000,5,1) # simulates data on 40k tests

# identify a subset of strong tests
m.1by1 = mash_1by1(mash_set_data(simdata$Bhat,simdata$Shat))
strong.subset = get_significant_results(m.1by1,0.05)

# identify a random subset of 5000 tests
random.subset = sample(1:nrow(simdata$Bhat),5000)
```

and create `random` and `strong` sets,

```{r}
data.temp = mash_set_data(simdata$Bhat[random.subset,],simdata$Shat[random.subset,])
Vhat = estimate_null_correlation(data.temp)
rm(data.temp)
data.random = mash_set_data(simdata$Bhat[random.subset,],simdata$Shat[random.subset,],V=Vhat)
data.strong = mash_set_data(simdata$Bhat[strong.subset,],simdata$Shat[strong.subset,], V=Vhat)
```

# FLASH analysis

**FIXME**

1. (we can review the code here and possibly implement `cov_flash` for future releases)
2. (For the time being we remove the additional decomposition on FLASH loading because there are issues needs to be figured out. @yuxin)

```{r}
library(flashr)
flash_pipeline = function(data, ...) {
  ## FIXME: need to adopt the current state-of-the art from Jason
  return(flash(data, ...))
}

cov_flash = function(data, subset = NULL, mode = 'EZ', save_model = NULL) {
  if(is.null(subset)) subset = 1:mashr:::n_effects(data)
  ## FIXME: is this reasonable?
  if (mode == 'EZ') b = data$Bhat / data$Shat
  else b = data$Bhat
  b.center = apply(b, 2, function(x) x - mean(x))
  fdata = flash_set_data(b.center)
  fmodel = flash_pipeline(fdata, greedy = TRUE, backfit = TRUE)
  flash_f = fmodel$ldf$f
  row.names(flash_f) = colnames(b)
  if (!is.null(save_model)) saveRDS(fmodel, save_model)
  U.flash = c(cov_from_factors(t(as.matrix(flash_f)), "Flash"),
      list("tFlash" = t(fmodel$fitted_values) %*% fmodel$fitted_values / nrow(b)))
  return(U.flash)
}
```

```{r}
U.f = cov_flash(data.strong)
```

# Finalize covariances

```{r}
U.pca = cov_pca(data.strong, 5)
U.ed = cov_ed(data.strong, c(U.f$U.flash, U.f$U.loading, U.pca))
U.c = cov_canonical(data.random)
```
    
## Fit mash model (estimate mixture proportions)

Now we fit mash to the random tests using both data-driven and canonical covariances. 
```{r}
m = mash(data.random, Ulist = c(U.ed,U.c), outputlevel = 1)
```

# Compute posterior summaries

Now we can compute posterior summaries etc for any subset of tests using the above mash fit. Here we do this for the `strong` tests.
```{r}
m2 = mash(data.strong, g=get_fitted_g(m), fixg=TRUE)
head(get_lfsr(m2))
```
[eqtl]: https://stephenslab.github.io/mashr/articles/eQTL_outline.html
