---
title: 'Introduction to mash: data-driven covariances'
author: "Matthew Stephens"
date: "May 30, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Goal

This vignette introduces the important topic of using  
data-driven covariance
matries in `mashr`. You should read [the introductory vignette](intro_mash.html) before this.

# Outline

Recall the four major steps in a mash analysis:
- Read in the data
- Set up the covariance matrices to be used
- Fit the model
- Extract posterior summaries and other quantities of interest

Here we pay more attention to Step 2, which can be split into
two steps:
- Set up the "canonical" covariance matrices.
- Set up the "data-driven" covariance matrices.

The first of these steps ("canonical" covariance matrices) is straightforward using `cov_canonical`, as illustrated in 
[the introductory vignette](intro_mash.html).

Setting up the data-driven matrices is slightly more complex,
and there are multiple possible approaches. This vignette illustrates
one approach.

First we simulate some data for illustration (see the [the introductory vignette](intro_mash.html) for more details):
```{r}
  library("mashr")
  set.seed(1)
  simdata = simple_sims(500,5,1)
```

# Data-driven covariances

Although the user is free to set up data driven covariance matrices
using any method that they would like, typically we suggest the following
three-step strategy:

1. locate some strong signals (e.g. by running a "condition-by-condition" using `mash_1by1`).
2. Use methods such as PCA or factor analysis on these top signals to compute some initial data-driven covariance matrices (e.g. using `cov_pca`)
3. Use these data-driven covariance matrices as initializations for the extreme deconvolution (ED) algorithm (using `cov_ed`), to get some refined data-driven covariance matrix estimates.

The ED step is helpful primarily 
because the second step will often estimate
the covariances of the observed data (Bhat) whereas what is required is the
covariances of the actual underlying effects (B), and this is what
ED estimates. However, ED can be quite sensitive to initialization,
and so the goal of the second step is to provide a good initialization.

## Step 1: select strong signals

The simplest way to do this is to run a condition-by-condition (1by1) analysis. Here we select the strong signals as those with lfsr<0.05 in any condition in the 1by1 analysis.
```{r}
  data = set_mash_data(simdata$Bhat, simdata$Shat)
  m.1by1 = mash_1by1(data)
  strong = get_significant_results(m.1by1,0.05)
```

## Step 2: Obtain initial data-driven covariance matrices

There are various approaches to this; here we just illustrate the simplest and quickest (but probably not the best), which is to use PCA.
Specifically here we use the function `cov_pca`
to produce covariance matrices based on the top 5 PCs of the strong signals.
The result is a list of 6 covariance matrices: one based on all 5 PCs, and the others each based on one PC.
```{r}
  U.pca = cov_pca(data,5,strong)
  names(U.pca)
```

## Step 3: Apply Extreme Deconvolution

The function `cov_ed` is used to apply the ED algorithm from a specified
initialization (here `U.pca`) and to a specified subset of signals.
```{r}
  U.ed = cov_ed(data, U.pca, strong)
```


# Run mash

After all this we are ready to run `mash` using the data driven matrices.
```{r}
  m.ed = mash(data, U.ed)
  get_loglik(m.ed)
```

From the fit we see that the fit using the data-driven covariances
is not as good as when we used 
the canonical covariances (which was -16120.32; from [the introductory vignette](intro_mash.html)). This is expected for this
simulation because the simulation actually used the canonical covariances!

In general we recommend running `mash` with both data-driven
and canonical covariances. You could do this by combining the data-driven
and canonical covariances as in this (unrun) code:

```{r}
  U.c = cov_canonical(data)  
  m = mash(data, c(U.c,U.ed))
```

For an example with simulations that do not follow the standard canonical matrices see [here](simulate_noncanon.html)
